---
title: "M8: Forecast Accuracy & Model Performance - Exercise"
author: "Luana Lima"
output: pdf_document
always_allow_html: true
editor_options: 
  chunk_output_type: inline
---

## Setting R code chunk options

First R code chunk is used for setting the options for all R code chunks. The choice echo=TRUE means both code and output will appear on report, include = FALSE neither code nor output is printed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), tidy=FALSE) 
```

## Loading packages and initializing

Second R code chunk is for loading packages. By setting message = FALSE, the code will appear but not the output.

```{r package, message=FALSE}
library(lubridate)
library(ggplot2)
library(forecast)  
#library(Kendall)
library(tseries)
#library(outliers)
library(tidyverse)
library(smooth)

#New package for M9 to assist with tables
#install.packages("kableExtra")
library(kableExtra)

```

## Understanding accuracy metrics

Compute metrics using the formulas below and then compare your results with accuracy().

$ME = \sum(Actual - Forecast)/n$
$MSE = \sum(Actual - Forecast)^2/n$ 
$RMSE = sqrt(MSE)$
$MAD = MAE = \sum|Actual - Forecast|/n$ 
$MPE = \sum((Actual - Forecast)/Actual)/n*100$
$MAPE = \sum|Actual - Forecast|/Actual/n*100$ 

```{r}
actual=c(107,125,115,118,108)
forecast=c(110,121,112,120,109)

#Exercise 1: compute MAD, MSE and MAPE
n=length(actual)

ME = sum(actual - forecast)/n
MSE = sum((actual - forecast)^2)/n
RMSE = sqrt(MSE)
MAE = sum(abs(actual - forecast))/n
MPE = sum((actual - forecast)/actual)/n*100
MAPE = sum(abs(actual - forecast)/actual)/n*100

manual_scores <- data.frame(ME,MSE,RMSE,MAE,MPE,MAPE)
manual_scores

#Exercise 2: Compute the forecast accuracy metrics you calculated in Ex 1 with the results from accuracy() function.
auto_scores <- as.data.frame(accuracy(forecast,actual))
auto_scores

```

## Importing data

Today we continue working with data from the Climate Change Knowledge Portal from the World Bank Group. More specifically historical rainfall and temperature averages for Brazil. You will find two new data files on folder "/Data/Raw/". One with rainfall named "pr_1901_2016_BRA.csv" and another with temperature named "tas_1901_2016_BRA.csv". The data span the period from 1901 to 2016 in monthly steps. You can download the data [here][<https://climateknowledgeportal.worldbank.org/download-data>]

Research question: Can you forecast temperature for the next two months?

```{r}
# Import both datasets using the read.csv function.
Rainfall_BR <- read.csv("./Data/pr_1901_2016_BRA.csv", stringsAsFactors = TRUE)

Temp_BR <- read.csv("./Data/tas_1901_2016_BRA.csv", stringsAsFactors = TRUE)

# Tidy the rainfall data sets.
#a Rename the column with Rainfall to get rid of the dots.
#b Note that on both data sets that is a column with the month name and average. Convert it to a Month column by separating the string, i.e., month in one column and the word "Average" on another. Hint: use the separate() function in the pipe.
#c Now you should have a column with Month. Use the paste0() function to paste month and year together and create a date object.
#d Select only the columns of interest: Date and rainfall

Rainfall_BR_processed <-
  Rainfall_BR %>% 
  rename( Rainfall = Rainfall....MM.) %>% 
  separate(Statistics,c("Null","Month","Null2")," ") %>% 
  mutate( Date = my(paste0(Month,"-",Year))) %>% 
  select(Date,Rainfall)

#Repeat for the temperature dataset.
Temp_BR_processed <-
  Temp_BR %>% 
  rename( Temperature_C = Temperature....Celsius.) %>% 
  separate(Statistics,c("Null","Month","Null2")," ") %>% 
  mutate( Date = my(paste0(Month,"-",Year))) %>% 
  select(Date,Temperature_C)  


#Join the temperature and rainfall into one tidy data frame with 3 columns: Date, Rainfall and Temperature.
BR_complete <- inner_join(Rainfall_BR_processed,Temp_BR_processed)
  
           
```

## Transforming data into time series object

```{r}
f_month <- month(first(BR_complete$Date))
f_year <- year(first(BR_complete$Date))

BR_Rain_ts <- ts(BR_complete$Rainfall,frequency = 12)
BR_Temp_ts <- ts(BR_complete$Temperature_C,frequency = 12)

```

## Decomposing the time series

```{r}
BR_Rain_Decomposed <- stl(BR_Rain_ts, s.window = "periodic")
plot(BR_Rain_Decomposed)

BR_Temp_Decomposed <- stl(BR_Temp_ts, s.window = "periodic")
plot(BR_Temp_Decomposed)

BR_Temp_ts_deseas <- seasadj(BR_Temp_Decomposed)

```

## Forecasting non-seasonal Temperature

Which models could be used to forecast the non-seasonal temperature series? When fitting a model it's important to do the backward-looking assessment. Check residuals to see if your model is well representing the historical data.

```{r}
#Exercise 3: Try fitting different model to the deseasonal series. Check teh residuals using the checkresiduals() function.

# Model 1: Naive on deseas data
NAIVE_deseas <- naive(BR_Temp_ts_deseas, h=12)
checkresiduals(NAIVE_deseas)

# Model 2:  ARIMA on deseasonal data
ARIMA_autofit <- auto.arima(BR_Temp_ts_deseas, max.D = 0, max.P = 0, max.Q = 0)
checkresiduals(ARIMA_autofit)

# Model 3:  Simple exponential smoothing on deseasonal data
SES_deseas=ses(y = BR_Temp_ts_deseas, h = 12, holdout = FALSE, silent = FALSE)  #alpha is NULL
checkresiduals(SES_deseas)

```

## Model Performance for forecasting 12 steps ahead

In the lesson we used function `accuracy()` to calculate performance measures like **ME:** Mean Error **RMSE:** Root Mean Squared Error **MAE:** Mean Absolute Error **MPE:** Mean Percentage Error **MAPE:** Mean Absolute Percentage Error **MASE:** Mean Absolute Scaled Error **ACF1:** Autocorrelation of errors at lag 1

```{r}

#Exercise 4: Apply function accuracy() to the forecasts generated with each model you fit for the non-seasonal data on Ex 3. 

#Model 1: Arithmetic mean
NAIVE_scores <- accuracy(NAIVE_deseas)  #store the performance metrics
plot(NAIVE_deseas)  #plot forecasts

# Model 2:  ARIMA 
#remember auto.arima does not call the forecast() internally so we need one more step
ARIMA_forecast <- forecast(object = ARIMA_autofit, h = 12)
ARIMA_scores <- accuracy(ARIMA_forecast)
plot(ARIMA_forecast)

# Model 3:  Simple exponential smoothing on deseasonal data
SES_scores <- accuracy(SES_deseas)  #store the performance metrics
plot(SES_deseas)  #

```

A better visualization of the results from `accuracy()` can be obtained by creating a data frame where rows correspond to models and columns to metrics. You can choose one metric to help you choose among models.

```{r}
# Exercise 5: create data frame that combines metrics for all the models you created on Ex 3.
deseas_scores <- as.data.frame(rbind(NAIVE_scores, ARIMA_scores, SES_scores))
row.names(deseas_scores) <- c("NAIVE", "ARIMA","SES")

# Exercise 6: Decide which model is the best fit by comparing the RMSE metric, i.e, choose model with lowest RMSE
best_model_index <- which.min(deseas_scores[,"RMSE"])
cat("The best model by RMSE is:", row.names(deseas_scores[best_model_index,]))                       
                            
```

## Visualization Challenge

```{r}
# Exercise 7: Generate a comparison table for your report using the kbl() function. 

kbl(deseas_scores, 
      caption = "Forecast Accuracy for Non-seasonal Temperature Data",
      digits = array(5,ncol(deseas_scores))) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  #highlight model with lowest RMSE
  kable_styling(latex_options="striped", stripe_index = which.min(seas_scores[,"RMSE"]))
```
